---
title: "Assigning Tidal Covariates to Harbor Seal Haul-out"
subtitle: ""
draft: true
author:
- name: Josh M. London
  affiliation: 1
address:
- code: 1
  address: Alaska Fisheries Science Center, NOAA, Seattle, Washington, USA 
  email: josh.london@noaa.gov
  orcid: orcid.org/0000-0002-3647-5046
date: "`r format(Sys.Date(), '%d %B, %Y')`"
disclaimer: >
  The scientific results and conclusions, as well as any views or opinions 
  expressed herein, are those of the author(s) and do not necessarily reflect 
  those of NOAA or the Department of Commerce.
abstract: >
  This document describes the process of assigning tidal covariates to Alaska
  harbor seal haul-out behavior derived from telemetry deployments. The data
  for haul-out behavior are pulled from the `pephaulout` R data package. The
  tidal calculations are determined via the XTide library which must be 
  installed along with the `tidextractr` package.
output: 
  uswebr::html_uswds:
    number_sections: FALSE
---

```{r global-options, include=FALSE}
  knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Introduction

```{r}
library(pephaulout)
library(tidextractr)
library(tidyverse)
```

## percent timelines data w/ locations

the `pephaulout` package has haulout data from all species deployments within
the PEP. Here, we'll filter out just the "Pv" species. We will also create
a `tide_dt` column that is set to the half-hour. We'll use this for our tidal
covariate prediction so it represents the mid-point of the hourly haul-out
behavior.

```{r}
data(tbl_percent_locs)

tbl_percent_locs <- tbl_percent_locs %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(species == "Pv") %>% 
  dplyr::arrange(speno, date_time) %>% 
  dplyr::mutate(tide_dt = date_time + lubridate::dhours(0.5)) 
tbl_percent_locs %>% glimpse()
```

## survey polygons w/ tide stations

Next, we need to associate all of the haul-out behavior locations with a
survey polygon (and, thus, the assigned tidal station for that polygon).

We'll get the survey polygons from the PEP PostgreSQL database.

```{r db-connect}
library(DBI)
con <- dbConnect(RPostgres::Postgres(), 
                 host = Sys.getenv("PEP_IP"),
                 user = keyringr::get_kc_account("pgpep_londonj"),
                 password = keyringr::decrypt_kc_pw("pgpep_londonj"),
                 dbname = "pep")
```

The `akpvpolys` object is an `sf` object that represents each of the survey
polygons. For those haul-out locations within a polygon, we can join easily
using the `st_intersets` function. However, there are a significant number
of locations outside of a polygon and we will want to link those to the 
nearest polygon. For this, we'll rely on the `nabor` package. This package
can only work with point data, so we'll create a `akpvpolys_cent` object
that represents the centroids of each survey polygon.

```{r}
library(sf)
akpvpolys <- st_read_db(con, query = "SELECT * FROM surv_pv_cst.geo_polys",
                       geom_column = 'geom') %>% 
  sf::st_transform(3571)
akpvpolys_cent <- akpvpolys %>% sf::st_centroid()

poly_idx <- tbl_percent_locs %>%
  sf::st_intersects(akpvpolys) %>% 
  purrr::map_int(1,.default = NA)

akpvpolys <- akpvpolys %>% 
  sf::st_set_geometry(NULL)

tbl_percent_locs <- tbl_percent_locs %>%
  dplyr::bind_cols(akpvpolys[poly_idx,c("polyid","station")])
```

At this point, `tbl_percent_locs` has `polyid` and `station` assigned for any
location that falls within a survey polygon. For the remaining locations, we'll
need to convert the point coordinates to a matrix and then relay on the
`nabor::knn()` function to determine the nearest neighbor. This also provides
us a distance calculation for the nearest neighbor.


```{r}
locs_na <- tbl_percent_locs %>% 
  dplyr::filter(is.na(polyid)) %>% 
  sf::st_coordinates()

polys_mat <- akpvpolys_cent %>% 
  sf::st_coordinates()

library(nabor)

locs_poly_nn <- knn(data=polys_mat, query=locs_na, k = 1) %>% 
  purrr::map(as_tibble) %>% dplyr::bind_cols() %>% 
  dplyr::rename(poly_idx = V1, nn_dist = V11) %>% 
  dplyr::mutate(polyid = akpvpolys_cent$polyid[poly_idx],
                station = akpvpolys_cent$station[poly_idx]) %>% 
  dplyr::select(-poly_idx)

locs_na <- tbl_percent_locs %>% 
  dplyr::filter(is.na(polyid)) %>% 
  dplyr::select(-c(polyid,station)) %>% 
  dplyr::bind_cols(locs_poly_nn)

tbl_percent_locs <- tbl_percent_locs %>% 
  dplyr::filter(!is.na(polyid)) %>% 
  dplyr::mutate(nn_dist = 0) %>% 
  rbind(locs_na) %>% 
  dplyr::arrange(speno,date_time)
```


## calculate tidal covariates

At this point, we're ready to do our tidal calculations. We'll rely on the 
`tidextractr` package (available from jmlondon/tidextracr on GitHub). This
package requires installation of the XTide command line tool.

We're going to take advantage of parallel processing using the `doFuture`
package and the `foreach` package. To help with this, we'll first nest our
data.

```{r}
tbl_percent_locs <- tbl_percent_locs %>% 
  dplyr::group_by(speno,species,sex,age) %>% tidyr::nest()
```


Now, we'll calcualte the tidal statistics we are interested in and bind those
back into `tbl_percent_locs`. This could take a while.

```{r}
library("doFuture")
library("foreach")
registerDoFuture()
plan(multiprocess)

get_tidestats <- function(tbl) {
  tide_height <- foreach(i = 1:nrow(tbl)) %do% {
    tidextractr::GetTideHeight(tbl$date_time[i], 
                               tbl$station[i])
  } %>% purrr::map_df(as_tibble) %>% 
    dplyr::rename(tide_height = value)
  
  near_low <- foreach(i = 1:nrow(tbl)) %do% {
    tidextractr::GetNearestLow(tbl$date_time[i], 
                               tbl$station[i])
  } %>% purrr::map_df(as_tibble) %>% 
    dplyr::rename(nlow_height = height,
                  nlow_time = time)
  
  dplyr::bind_cols(tbl,tide_height,near_low) %>% 
    mutate(minutes_from_low = difftime(date_time,nlow_time) %>% 
             as.numeric(units = "mins"))
}

library(tictoc)

tic("get tidestats")
tbl_percent_locs <- tbl_percent_locs %>% 
  dplyr::mutate(data = foreach(i = 1:nrow(.)) %dopar% {
    get_tidestats(data[[i]])
  })
toc()
```

```{r}
usethis::use_data(tbl_percent_locs)
```



