---
title: "Assigning Tidal Covariates to Harbor Seal Haul-out"
subtitle: ""
draft: true
author:
- name: Josh M. London
  affiliation: 1
address:
- code: 1
  address: Alaska Fisheries Science Center, NOAA, Seattle, Washington, USA 
  email: josh.london@noaa.gov
  orcid: orcid.org/0000-0002-3647-5046
date: "`r format(Sys.Date(), '%d %B, %Y')`"
disclaimer: >
  The scientific results and conclusions, as well as any views or opinions 
  expressed herein, are those of the author(s) and do not necessarily reflect 
  those of NOAA or the Department of Commerce.
abstract: >
  This document describes the process of assigning tidal covariates to Alaska
  harbor seal haul-out behavior derived from telemetry deployments. The data
  for haul-out behavior are pulled from the `pephaulout` R data package. The
  tidal calculations are determined via the XTide library which must be 
  installed along with the `tidextractr` package (available on Github).
output: 
  uswebr::html_uswds:
    number_sections: FALSE
vignette: >
  %\VignetteIndexEntry{Assigning Tidal Covariates to Harbor Seal Haul-out}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r global-options, include=FALSE}
  knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Introduction

```{r}
library(pephaulout)
library(tidextractr)
library(tidyverse)
library(furrr)
```

## percent timelines data w/ locations

the `pephaulout` package has haulout data from all species deployments within
the PEP. Here, we'll filter out just the "Pv" species. We will also create
a `tide_dt` column that is set to the half-hour. We'll use this for our tidal
covariate prediction so it represents the mid-point of the hourly haul-out
behavior.

```{r}
data(tbl_percent_locs)

tbl_percent_locs <- tbl_percent_locs %>% 
  dplyr::ungroup() %>% 
  dplyr::filter(species == "Pv") %>% 
  dplyr::arrange(speno, date_time) %>% 
  dplyr::mutate(tide_dt = date_time + lubridate::dhours(0.5)) %>% 
  sf::st_transform(3338)
tbl_percent_locs %>% glimpse()
```

## survey polygons w/ tide stations

Next, we need to associate all of the haul-out behavior locations with a
survey polygon (and, thus, the assigned tidal station for that polygon).

We'll get the survey polygons from the PEP PostgreSQL database.

```{r db-connect}
library(DBI)
con <- dbConnect(RPostgres::Postgres(), 
                 host = Sys.getenv("PEP_IP"),
                 user = keyringr::get_kc_account("pgpep_londonj"),
                 password = keyringr::decrypt_kc_pw("pgpep_londonj"),
                 dbname = "pep")
```

The `akpvpolys` object is an `sf` object that represents each of the survey
polygons. For those haul-out locations within a polygon, we can join easily
using the `st_intersets` function. However, there are a significant number
of locations outside of a polygon and we will want to link those to the 
nearest polygon. For this, we'll rely on the `nabor` package. This package
can only work with point data, so we'll create a `akpvpolys_cent` object
that represents the centroids of each survey polygon.

```{r}
library(sf)
akpvpolys <- st_read(con, 
                     query = "SELECT * FROM surv_pv_cst.geo_polys where iliamna != 'Y'") %>% 
  sf::st_transform(3338)
akpvpolys_cent <- akpvpolys %>% sf::st_centroid()

poly_idx <- tbl_percent_locs %>%
  sf::st_intersects(akpvpolys) %>% 
  purrr::map_int(1,.default = NA)

akpvpolys <- akpvpolys %>% 
  sf::st_set_geometry(NULL)

tbl_percent_locs <- tbl_percent_locs %>%
  dplyr::bind_cols(akpvpolys[poly_idx,c("polyid","station")])
```

At this point, `tbl_percent_locs` has `polyid` and `station` assigned for any
location that falls within a survey polygon (polygons associated with Iliamna
Lake are not included). Of these, we will want to separate out glacial fjord
habitats as those will not have tidal covariates associated with them. The
glacial polygons have a non NULL value in the `glacial_name` field. But, since
we do want records that are nearest to glacial habitat also associated with
those polygons, we will wait to separate out the glacial polygons until after
the nearest neighbors are assigned. For the remaining locations, we'll need to
convert the point coordinates to a matrix and then relay on the `nabor::knn()`
function to determine the nearest neighbor. This also provides us a distance
calculation for the nearest neighbor.

```{r}
glacial_polys <- st_read(con, query = "SELECT * FROM surv_pv_cst.geo_polys where glacier_name IS NOT NULL")
```

Now that we have a list of our glacial fjord polygons, we can proceed with
assignment of nearest neighbors.

```{r}
locs_na <- tbl_percent_locs %>% 
  dplyr::filter(is.na(polyid)) %>% 
  sf::st_coordinates()

polys_mat <- akpvpolys_cent %>% 
  sf::st_coordinates()

library(nabor)

locs_poly_nn <- knn(data=polys_mat, query=locs_na, k = 1) %>% 
  purrr::map(as_tibble) %>% dplyr::bind_cols() %>% 
  dplyr::rename(poly_idx = V1, nn_dist = V11) %>% 
  dplyr::mutate(polyid = akpvpolys_cent$polyid[poly_idx],
                station = akpvpolys_cent$station[poly_idx]) %>% 
  dplyr::select(-poly_idx)

locs_na <- tbl_percent_locs %>% 
  dplyr::filter(is.na(polyid)) %>% 
  dplyr::select(-c(polyid,station)) %>% 
  dplyr::bind_cols(locs_poly_nn)

tbl_percent_locs <- tbl_percent_locs %>% 
  dplyr::filter(!is.na(polyid)) %>% 
  dplyr::mutate(nn_dist = 0) %>% 
  rbind(locs_na) %>% 
  dplyr::arrange(speno,date_time)

tbl_percent_glacial <- tbl_percent_locs %>% 
  dplyr::filter(polyid %in% glacial_polys$polyid) 

tbl_percent_locs <- tbl_percent_locs %>% 
  dplyr::filter(!polyid %in% glacial_polys$polyid)
```

All of our haul-out behavior records are now associated with a tidal station.
Let's take a moment and examine the distribution of distances from the haul-out
behavior to the nearest survey unit.

```{r}
tbl_percent_locs %>% dplyr::filter(nn_dist > 50000) %>% 
  ggplot(aes(nn_dist/1000)) + geom_density(fill = "#003366", size = 0) +
  xlab("distance (km)") +
  ggtitle("Distribution of nearest distance between haul-out location and survey units",
          subtitle = "only showing distances > 50 km")
```

The majority of the records are within 250 km of a survey unit (and, thus, a
tidal prediction location). Distances greater than 250 km likely represent
unrealistic locations.

Lets take a look at a map of those extreme records

```{r}
library(mapview)
tbl_percent_locs %>% dplyr::filter(nn_dist > 250000) %>% 
  mapview()
```

Given this map, it seems prudent to remove those records from the dataset and
not include them as these locations are not realistic for harbor seals.

```{r}
tbl_percent_locs <- tbl_percent_locs %>% dplyr::filter(nn_dist <= 250000)
```

Now, let's take a look at the remaining locations (note, we have to sample
in order to display w/out overplotting too much)

```{r}
library(nPacMaps)
ak_base <- nPacMaps::alaska()

locs <- tbl_percent_locs %>% dplyr::sample_frac(0.025) 

ggplot() +
  geom_sf(data = ak_base, size = 0, fill = "grey70") +
  geom_sf(data = locs, size = 0.1, alpha = 0.2) +
  scale_x_continuous(breaks = c(180, -160, -140)) +
  ggtitle("Spatial distribution of haul-out behavior records",
          subtitle = "note: only 2.5% of data shown")
```

Even some of these locations seem unrealistic (well inland). Downstream
analysis of data should consider additional censoring or downweighting of
records based on the `nn_dist` attribute.

## calculate tidal covariates

At this point, we're ready to do our tidal calculations. We'll rely on the 
`tidextractr` package (available from jmlondon/tidextracr on GitHub). This
package requires installation of the XTide command line tool.

We're going to take advantage of parallel processing using the `furrr`
package. To help with this, we'll first nest our data.

```{r}
tbl_percent_locs <- tbl_percent_locs %>% 
  dplyr::group_by(speno,species,sex,age) %>% tidyr::nest()
```


Now, we'll calcualte the tidal statistics we are interested in and bind those
back into `tbl_percent_locs`. This could take a while, but we will speed things
up a bit by using the `furrr` package for parallel processing.

```{r}

create_tidestats <- function(tbl_data) {
  get_tide_height <- purrr::possibly(tidextractr::GetTideHeight, 
                                     otherwise = NA, quiet = FALSE)
  get_nearest <- purrr::possibly(tidextractr::GetNearestLow, 
                                 otherwise = NA, quiet = FALSE)
  
  tbl_data <- tbl_data %>%
    dplyr::mutate(
      tide_height = purrr::map2_dbl(tide_dt, station,
                                    ~ get_tide_height(.x, .y)),
      near_low = purrr::map2(tide_dt, station,
                             ~ get_nearest(.x, .y)),
      near_low_height = purrr::map_dbl(near_low, "height"),
      near_low_time = purrr::map(near_low, "time") %>%
        purrr::reduce(c)
    ) %>% 
    dplyr::filter(inherits(near_low_time,"POSIXct")) %>%
    dplyr::mutate(
      minutes_from_low = purrr::map2(tide_dt, near_low_time, 
                                     ~ difftime(.x, .y, units = "mins")) %>%
        as.numeric()
    ) %>% 
    dplyr::select(-(near_low))
}

plan(multiprocess)

tbl_percent_locs <- tbl_percent_locs %>% 
  dplyr::mutate(data = furrr::future_map(data, create_tidestats)) 

```

The last step is to create our final `tbl_percent_locs` and `tbl_percent_glacial`
data objects.

```{r}
tbl_percent_locs <- tbl_percent_locs %>% tidyr::unnest()
usethis::use_data(tbl_percent_locs, overwrite = TRUE)
usethis::use_data(tbl_percent_glacial, overwrite = TRUE)
```
